PubCrawl: A Publication Search Engine

### 2012-05-18 report/paper.tex
Incorporate Manasi's writeup about the user study

We implemented two ranking functions. The first only uses the full text score
and PageRank. The second uses a rank join--inspired algorithm to combine lists
of hits sorted by each of title, abstract, and full text scores and PageRank.
Briefly, the algorithm maintains the top $k$ hits seen so far. In each
iteration, we pop the maximum element off of the next list in round-robin order
and hash it. When we encounter a duplicate, we update the combined score and
its position in the top $k$. The algorithm terminates when the maximum possible
score of unseen tuples is less than the score of the $k$th hit.

To tune weights for the four factors, we obtained information about important
papers on 6 topics from experts in the respective areas. We then manually
picked weights that gave good results for all 6 topics.

@@

To evaluate the quality of PubCrawl results, we performed a preliminary user study. In this study, we compared the quality of PubCrawl results with those from Google Scholar, Microsoft Academic Search, ACM Digital Library and \citeseer.

@@

For 3 topics, we performed searches on Google Scholar, Microsoft Academic Search, ACM Digital Library, CiteSeer and PubCrawl. We then obtained lists of the top 10 ACM publication results from these search engines. We provided these lists to our experts. The experts rated each of the 10 results as Highly Relevant, Moderately Relevant and Not relevant. They also ranked the five lists in order of decreasing quality. Table \ref{} shows the results for the 3 test topics.

We find that PubCrawl consistently does as well as CiteSeer and ACM DL.

\subsection{Challenges}
The main challenges with evaluating the user study were (1) Ground truth wasn't always clearly known, (2) Ground-truth is subjective, (3) since the data only included ACM papers, we had to discard a large number of highly relevant but non-ACM publications.

\section{Results}

User study results.

Some time comparisons.

@@

>>>>

We implemented two ranking functions. The first only uses the full text score
and PageRank. The second uses an algorithm based on RankJoin
\cite{DBLP:conf/vldb/IlyasAE03} to combine lists of hits sorted by each of
title, abstract, and full text scores and PageRank. The algorithm maintains the
top $k$ hits seen so far. In each iteration, we pop the maximum element off of
the next list in round-robin order and hash it. When we encounter a duplicate,
we update the combined score and its position in the top $k$. The algorithm
terminates when the maximum possible score of unseen tuples is less than the
score of the $k$th hit.

@@

To evaluate the quality of PubCrawl results, we performed a preliminary user
study comparing the quality of PubCrawl results with those from Google Scholar,
Microsoft Academic Search, ACM Digital Library and CiteSeer.

@@

We obtained information about important papers on 6 topics from experts in the
respective areas. We then manually picked weights, given in Table \ref{weights}
that gave good results for all 6 topics.

\begin{table}[bt]
  \begin{center}
    \begin{tabular}{ll}
      Parameter & Weight\\
      \hline
      Title & 0.1\\
      Abstract & 0\\
      Document & 0.5\\
      PageRank & 1\\
    \end{tabular}
  \end{center}
  \caption{Manually optimized parameter settings for PubCrawl user study}
  \label{weights}
\end{table}

For 4 topics, we performed searches on the engines, generating lists of the top
10 ACM publications from each. We provided these lists to our experts who rated
each of the 10 results as Highly Relevant, Moderately Relevant or Not relevant.
They also ranked the five lists in order of decreasing quality.

Figure \ref{relevance1} shows the distribution of results grouped by ranks
assigned by users. PubCrawl produces a large number of moderately relevant
document but a low number both of highly relevant and not relevant results.
While we would ideally like to maximize the former, the low number of
irrelevant results suggests are method has a lower false positive rate.

Figures \ref{relevance2} shows the relevance ratio across all search engines.
Relevance ratio is defined as $(H + M) / N$, where $H$ is the number of Highly
Relevant results, $M$ is the number of Moderately Relevant results, and $N$ is
the number of Not relevant results. PubCrawl has relevance ratio comparable to
Google Scholar and Microsoft Academic Search. CiteSeer, on the other hand, has
very poor relevance score since it produces a large number of irrelevant
results. ACM Digital Library has the largest number of relevant documents in
its results.

In Figure \ref{relevance3}, we show a total quality score which we define as $H
- N + .1 M$. We choose this metric to highly rank relevant results, penalize
for non-relevant results and factor in moderately relevant results.

Finally, in Figure \ref{rank1}, we show the average rank assigned to an engine
based on the ranks assigned to each engine by our experts. A higher score
(scale of 5) indicates a better rank. On average, PubCrawl ranked as high as
the other engines except Google Scholar.

\begin{figure}[p]
\centering
\includegraphics[width=5in]{images/Relevance1}
\caption{Relevance of search results grouped as Highly Relevant, Moderately
  Relevant and Not Relevant.}
\label{relevance1}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[width=5in]{images/Relevance2}
\caption{Relevance ratio vs. Search Engines.}
\label{relevance2}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[width=5in]{images/Relevance3}
\caption{Quality Score.}
\label{relevance3}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[width=5in]{images/Rank}
\caption{Rank order of search engines (Higher score indicates better rank; scale of 5).}
\label{rank1}
\end{figure}

\subsection{Discussion}

Our preliminary user study only involved 4 topics and hence a larger study must
be undertaken to rigorously study the quality of results. We also observed
that: (1) quality of results depends heavily on the breadth of the search term
(a narrow search term produces better results), (2) ground truth wasn't always
clearly known, (3) ground-truth is subjective, (4) we had to discard a large
number of highly relevant but non-ACM publications. Furthermore, because we did
not have data for non-ACM papers citing ACM papers, our PageRank algorithm was
weaker compared to other search engines. However, our preliminary user study
shows that PubCrawl can find relevant documents and effectively trade off
PageRank and text indexing scores.

@@

In this work, we built a publication search engine based on the PageRank
algorithm and the Lucene text index. We adapted PageRank for the specific
properties of the citation graph and indexed various paper attributes such as
title, abstract and full text. We then explored multiple ways of combining the
text indexing scores with PageRank. We observed that it is very difficult to
find a scoring function that works well for all search queries. We also
realized that combining scores from orthogonal metrics like PageRank and text
indexing may not be the best method to obtain an overall ranking. Our user
study demonstrates that PubCrawl has a low false positive rate and produces a
large number of moderately relevant results. However, it produces a low number
of highly relevant results and there is potential to extend this work by
finding better ranking functions.
